---
title: "data.table cheat sheet"
output: html_notebook
---

```{r}
library(data.table)
```

## Create a data.table

```{r}
data.table(a = c(1, 2), b = c("a", "b"))
```

```{r}
dt <- mtcars
setDT(dt)
dt
```

```{r}
dt <- as.data.table(dt)
dt
```

## Subset rows using i

```{r}
dt[1:2, ]
```

```{r}
dt[mpg > 20, ]
```

```{r}
dt[mpg %between% c(20, 25), ]
```

## Manipulate columns with j

### Extract

```{r}
dt[, c(2)]
```

```{r}
dt2 <- dt[, .(mpg, cyl)]
dt2
```

### Summarize

```{r}
dt2[, .(x = sum(mpg))]
```

```{r}
dt2[, c := 1 + 2]
dt2
```

```{r}
dt2[mpg == 21.0, d := 1 + 2]
dt2
```

```{r}
dt2[, `:=`(e = 1, f = 2)]
dt2
```

### Delete column

```{r}
dt2[, `:=`(c = NULL, d = NULL, e = NULL, f = NULL)]
dt2
```

### Convert column type

```{r}
dt2[, mc := mpg + cyl]
dt2[, mc := as.integer(mc)]
dt2
```

## Group according to by

```{r}
dt2[, .(c = sum(mpg)), by = cyl]
```

```{r}
dt2[, .(c = sum(mpg)), keyby = cyl] # sorted
```

### Common grouped operations

```{r}
dt2[, c := mean(mpg), by = cyl]
dt2
```

```{r}
dt2[, `:=`(mc = NULL, c = NULL)]
dt2
```

```{r}
dt2[, .SD[1], by = cyl] # 그룹별 첫번째 줄
```

```{r}
dt2[, .SD[.N], by = cyl] # 그룹별 마지막 줄
```

## Chaining

```{r}
dt2[, mc := mpg + cyl][, .(s = mean(mc)), by = cyl]
```

## Functions for data.tables

### Reorder

```{r}
setorder(dt2, cyl, -mpg)
dt2
```

### Unique rows

```{r}
unique(dt2, by = c("cyl"))
```

```{r}
uniqueN(dt2, by = c("cyl"))
```

### Rename columns

```{r}
setnames(dt2, c("mpg", "cyl"), c("MPG", "CYL"))
dt2
```

## Combine data.tables

### Join

```{r}
dt_a <- data.table(
  a = c(1, 2, 3),
  b = c("c", "a", "b")
)
dt_a

dt_b <- data.table(
  x = c(3, 2, 1),
  y = c("b", "c", "a")
)
dt_b
```

```{r}
dt_a[dt_b, on = .(b = y)]
```

```{r}
dt_b[dt_a, on = .(y = b)]
```

```{r}
dt_a <- data.table(
  a = c(1, 2, 3),
  b = c("c", "a", "b"),
  c = c(7, 5, 6)
)
dt_a

dt_b <- data.table(
  x = c(3, 2, 1),
  y = c("b", "c", "a"),
  z = c(4, 5, 8)
)
dt_b
```

```{r}
dt_a[dt_b, on = .(b = y, c > z)]
```

### Rolling join

```{r}
dt_a <- data.table(
  a = c(1, 2, 3, 1, 2),
  id = c("A", "A", "A", "B", "B"),
  date = c(
    as.Date("2010-01-01"),
    as.Date("2012-01-01"),
    as.Date("2014-01-01"),
    as.Date("2010-01-01"),
    as.Date("2012-01-01")
  )
)
dt_a
dt_b <- data.table(
  b = c(1, 1),
  id = c("A", "B"),
  date = c(as.Date("2013-01-01"), as.Date("2013-01-01"))
)
dt_b
```

```{r}
dt_a[dt_b, on = .(id = id, date = date), roll = TRUE]
```

```{r}
dt_a[dt_b, on = .(id = id, date = date), roll = -Inf]
```

### Bind

```{r}
dt_a <- dt2[1:5, ]
dt_b <- dt2[6:10, ]
```

```{r}
rbind(dt_a, dt_b)
```

```{r}
a <- cbind(dt_a, dt_b) # 주의 동일한 이름이 존재
a
```

## Reshape a data.table

```{r}
dt <- data.table(
  id = c("A", "A", "B", "B"),
  y = c("x", "z", "x", "z"),
  a = c(1, 2, 1, 2),
  b = c(3, 4, 3, 4)
)
dt
dt2 <- dcast(dt, id ~ y, value.var = c("a", "b"))
dt2
```

```{r}
melt(dt2,
  id.vars = c("id"),
  measure.vars = patterns("^a", "^b"),
  variable.name = "y",
  value.name = c("a", "b")
)
```

## Apply function to cols

### Apply a function to multiple columns

```{r}
dt <- as.data.table(mtcars[1:10, 1:2])
dt
```

```{r}
dt[, lapply(.SD, mean), .SDcols = c("mpg", "cyl")]
```

```{r}
cols <- c("mpg", "cyl")
dt[, paste0(cols, "_m") := lapply(.SD, mean), .SDcols = cols]
dt
```

## Sequential rows

### Row ids

```{r}
dt[, c := 1:.N, by = cyl]
dt
```

### Lag & Lead

```{r}
dt <- as.data.table(mtcars[1:10, 1:2])
dt2 <- dt[order(cyl), ]
dt2[, c := shift(mpg, 1), by = cyl]
dt2
```

```{r}
dt <- as.data.table(mtcars[1:10, 1:2])
dt2 <- dt[order(cyl), ]
dt2[, c := shift(mpg, 1, type = "lead"), by = cyl]
dt2
```
